---
title: "Affix polyfunctionality in French deverbal nominalizations"
author: "Justine Salvadori, Richard Huyghe"
output:
  html_document:
    number_sections: true
    df_print: paged
---

# Packages and datasets

```{r Working directory}
setwd("XXX")
```

```{r Packages, message=FALSE, warning=FALSE}
library(readxl)
library(dplyr)
library(fuzzySim)
library(tidyverse)
library(pheatmap)
library(cooccur)
library(data.table)
library(vegan)
library(knitr)
library(cluster)
library(ggpubr)
library(MeanRarity)
```

```{r Datasets}
## Original data set
data_complete <- read_excel("data_complete.xlsx", sheet = 1)
data_complete

## Adapted data set (without functions realized only once per suffix)
data <- data_complete %>% 
        group_by(SUFFIX, COMBINED_TYPE) %>% 
        filter(n() > 1) %>% 
        ungroup()
data

## Data set for frequency of functions
data_freq <- read_excel("data_freq.xlsx", sheet = 1)
data_freq
```


# Distribution of functions among suffixes

## Suffix heterogeneity and functional variety

```{r Distribution of functions}
## Preparation of the df
p_a_complete <- splist2presabs(data, 
                               sites.col = "COMBINED_TYPE", 
                               sp.col = "SUFFIX", 
                               keep.n = FALSE)
names(p_a_complete)[names(p_a_complete) == 'in.'] <- 'in' 
p_a_complete <- p_a_complete %>% 
                remove_rownames %>% 
                column_to_rownames(var = "COMBINED_TYPE")
colnames(p_a_complete) <- paste("-", colnames(p_a_complete), sep="")
p_a_complete <- p_a_complete[ ,order(colSums(-p_a_complete))]
p_a_complete <- p_a_complete[order(rowSums(-p_a_complete)), ]
p_a_complete

## Plot => Figure 1 
pheatmap(p_a_complete,
         legend = F,
         fontsize = 4,fontsize_row = 4,
         cellwidth = 4, cellheight = 4,
         color = c(NA,"gray39"), border_color = "white", 
         las = 1, 
         cluster_rows = FALSE, cluster_cols = FALSE)
```

## Associations between functions

### Ontological functions

```{r Ontological functions}
## Preparation of the df
p_a_onto <- splist2presabs(data, 
                           sites.col = "ONTO_TYPE", 
                           sp.col = "SUFFIX", 
                           keep.n = FALSE)
p_a_onto <- p_a_onto %>% 
            remove_rownames %>% 
            column_to_rownames(var = "ONTO_TYPE")
p_a_onto

## Summary of significant results
res_coocc_onto <- cooccur(p_a_onto, spp_names = TRUE, thresh = TRUE)
summary(res_coocc_onto)

## Table of results
prob.table(res_coocc_onto)

## Summary of positive, negative and random interactions for each function
pair.attributes(res_coocc_onto)

## Plot
plot(res_coocc_onto) + ggtitle("") + theme(legend.position="bottom",
                                           legend.text = element_text(size=12))
```

### Relational functions

```{r Relational functions}
## Preparation of the df
p_a_relation <- splist2presabs(data, 
                               sites.col = "RELATIONAL_TYPE", 
                               sp.col = "SUFFIX", 
                               keep.n = FALSE)
p_a_relation <- p_a_relation %>% 
                remove_rownames %>% 
                column_to_rownames(var = "RELATIONAL_TYPE")
p_a_relation

## Summary of significant results
res_coocc_rela <- cooccur(p_a_relation, spp_names = TRUE, thresh = TRUE)
summary(res_coocc_rela)

## Table of results
prob.table(res_coocc_rela)

## Summary of positive, negative and random interactions for each function
pair.attributes(res_coocc_rela)

## Plot
plot(res_coocc_rela) + ggtitle("") + theme(legend.position="bottom",
                                           legend.text = element_text(size=12))
```

### Combined functions

```{r Combined functions}
## Preparation of the dataframe
p_a_complete <- splist2presabs(data, 
                               sites.col = "COMBINED_TYPE", 
                               sp.col = "SUFFIX", 
                               keep.n = FALSE)
p_a_complete <- p_a_complete %>% 
                remove_rownames %>% 
                column_to_rownames(var="COMBINED_TYPE")
p_a_complete

## Summary of significant results
res_coocc_complete <- cooccur(p_a_complete, spp_names = TRUE, thresh = TRUE)
summary(res_coocc_complete)

## Table of results
prob.table(res_coocc_complete)

## Summary of positive, negative and random interactions for each function
pair.attributes(res_coocc_complete)

## Plot ==> Figure 2
plot(res_coocc_complete) + ggtitle("") + theme(legend.position="bottom",
                                               legend.text = element_text(size=12))

```


## Hierarchy of functions

```{r Secondary rate}
## Secondary functions
sec_functions <-  data %>%
                  group_by(NOUN, SUFFIX, VERB) %>% 
                  mutate(NB_SENSE_POLYSEMY = n()) %>%
                  ungroup() %>%
                  mutate(POLYSEMY = if_else(NB_SENSE_POLYSEMY > 1, "y", "n")) %>% 
                  select(SUFFIX, COMBINED_TYPE, POLYSEMY) %>%
                  distinct(.keep_all = TRUE) %>%
                  group_by(SUFFIX, COMBINED_TYPE) %>%
                  filter(n() == 1, POLYSEMY=='y') %>%
                  select(-POLYSEMY) %>%
                  ungroup() %>%
                  group_by(COMBINED_TYPE) %>%
                  summarise(NB_SUFF_SEC = n_distinct(SUFFIX))
sec_functions

## Number of suffixes that realize each secondary function
complete_suff <- data %>%
                 group_by(COMBINED_TYPE) %>%
                 summarise(NB_SUFF_TOTAL = n_distinct(SUFFIX),
                 NB_MEANINGS = n()) %>%
                 arrange(-NB_SUFF_TOTAL)
complete_suff

## Rate of secondary realization 
sec_rate <- merge(sec_functions, complete_suff, by = 'COMBINED_TYPE')
sec_rate$SEC_RATE <- sec_rate$NB_SUFF_SEC / sec_rate$NB_SUFF_TOTAL 
sec_rate

## Plot ==> Figure 4
sec_rate %>%
  arrange(SEC_RATE) %>% 
  mutate(COMBINED_TYPE = factor(COMBINED_TYPE, levels=COMBINED_TYPE)) %>% 
  ggplot(aes(x = COMBINED_TYPE, y = SEC_RATE)) +
  geom_segment(aes(xend = COMBINED_TYPE, yend = 0)) +
  geom_point(shape = 15) +
  coord_flip() +
  xlab("Combined functions \n ") +
  ylab("\n Secondarity rate") + 
  expand_limits(y = 1.02) +
  scale_y_continuous(expand = c(0,0), breaks = seq(0, 1, 0.1)) +
  theme_bw() + 
  theme(text = element_text(size = 12),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"))
```


## Polyfunctionality and rivalry

```{r Clustering}
## Preparation of the df
p_a_complete <- splist2presabs(data, 
                               sites.col = "COMBINED_TYPE", 
                               sp.col = "SUFFIX", 
                               keep.n = FALSE)
names(p_a_complete)[names(p_a_complete) == 'in.'] <- 'in' 
p_a_complete <- p_a_complete %>% 
                remove_rownames %>% 
                column_to_rownames(var="COMBINED_TYPE")
colnames(p_a_complete) <- paste("-", colnames(p_a_complete), sep="")
p_a_complete <- t(p_a_complete)

## Distance matrix
distance_suff <- vegdist(p_a_complete, method="bray")  
as.data.frame(as.matrix(distance_suff))

## Similarity matrix
simil_suff <- 1-as.matrix(distance_suff) 
as.data.frame(as.matrix(simil_suff))

## Clustering (method: "average")
cluster_average <- hclust(distance_suff, method = "average")

## Plot (heatmap + dendrogram) ==> Figure 5 
pheatmap(simil_suff,
         scale = "none",
         clustering_method = "average",
         clustering_distance_cols = distance_suff,
         clustering_distance_rows = distance_suff)

## Optimal number of clusters (matrix correlation statistic)
#### Code from Borcard, D., Gillet, F., & Legendre, P. (2018). Numerical Ecology with R. Springer International Publishing. https://doi.org/10.1007/978-3-319-71404-2
hc <- cluster_average
kt <- data.frame(k = 1:nrow(p_a_complete), r = 0)
grpdist <- function(X)
{
	require(cluster)
	veg <- as.data.frame(as.factor(X))
	distgr <- daisy(veg, "gower")
	distgr
}
for (i in 2:(nrow(p_a_complete) - 1)) {
  gr <- cutree(hc, i)
  distgr <- grpdist(gr)
  mt <- cor(distance_suff, distgr, method = "pearson") 
  kt[i, 2] <- mt
}
k.best <- which.max(kt$r) 
plot(
    kt$k,
    kt$r,
    type = "h",
    main = "Matrix correlation-optimal number of clusters",
    xlab = "k (number of clusters)",
    ylab = "Pearson's correlation"
) 
axis(
  1,
  k.best,
  paste("optimum", k.best, sep = "\n"), 
  col = "red",
  font = 2,
  col.axis = "red"
) 
points(k.best,
  max(kt$r), 
  pch = 16, 
  col = "red", 
  cex = 1.5
)
```

# Frequency of functions 

## Samplying and analysis

```{r Sampling}
## List of functions previously identified for -ade, -ment and -ure (+ -on)
ade_fct <- data %>%
           filter(SUFFIX == "ade") %>%
           select(COMBINED_TYPE) %>%
           distinct(.keep_all = TRUE)

ment_fct <- data %>%
            filter(SUFFIX == "ment") %>%
            select(COMBINED_TYPE) %>%
            distinct(.keep_all = TRUE)

ure_fct <- data %>%
           filter(SUFFIX == "ure") %>%
           select(COMBINED_TYPE) %>%
           distinct(.keep_all = TRUE)

on_fct <- data %>%
          filter(SUFFIX == "on") %>%
          select(COMBINED_TYPE) %>%
          distinct(.keep_all = TRUE)

## Relevant meanings
keep <- data_freq %>%
        select(SUFFIX, NOUN, VERB, COMBINED_TYPE) %>%
        mutate(EXISTING_FCT = case_when(
                (SUFFIX == "ade" & COMBINED_TYPE %in% ade_fct$COMBINED_TYPE) ~ "y",
                (SUFFIX == "ment" & COMBINED_TYPE %in% ment_fct$COMBINED_TYPE) ~ "y",
                (SUFFIX == "ure" & COMBINED_TYPE %in% ure_fct$COMBINED_TYPE) ~ "y",
                (SUFFIX == "on" & COMBINED_TYPE %in% on_fct$COMBINED_TYPE) ~ "y",
                 TRUE ~ "n")) %>%
        group_by(SUFFIX, COMBINED_TYPE) %>%
        mutate(COUNT = n()) %>%
        ungroup() %>%
        filter(EXISTING_FCT == "y" | COUNT > 1)
keep
```

## Polyfunctionality and diversity

```{r Frequencies}
## Prepare the df
plot_freq <- keep %>% filter(SUFFIX != "on") 
plot_freq$SUFFIX <- sub("^","-", plot_freq$SUFFIX)
plot_freq <- plot_freq %>%
             distinct(SUFFIX, COMBINED_TYPE, COUNT, .keep_all = FALSE) %>%
             arrange(SUFFIX, COMBINED_TYPE)
plot_freq

## Plot - Balloon ==> Figure 6
balloon <- ggplot(plot_freq, 
                  aes(x = SUFFIX, y = COMBINED_TYPE)) 
balloon + geom_point(aes(size = COUNT),
                     shape = 21,
                     fill = "darkgray",
                     color = "black") + 
          scale_y_discrete(limits = rev) +
          scale_x_discrete(position = "top") +
          theme_bw() + 
          theme(text = element_text(size = 15),
                panel.border = element_blank(),
                axis.title.x = element_blank(),
                axis.title.y = element_blank()) +
          labs(size = "Frequency")
```

```{r Hill-Shannon diversity}
## Abundance table
hill_abund <- keep %>% filter(SUFFIX != "on")
hill_abund <- splist2presabs(hill_abund, 
                             sites.col = "SUFFIX", 
                             sp.col = "COMBINED_TYPE", 
                             keep.n = TRUE)
hill_abund <- hill_abund %>%
              remove_rownames %>% 
              column_to_rownames(var="SUFFIX")
hill_abund

## Hill-Shannon diversity: -ade
ade_abund <- as.numeric(as.vector(hill_abund[1,]))
rarity(ade_abund, 0)

## Hill-Shannon diversity: -ment
ment_abund <- as.numeric(as.vector(hill_abund[2,]))
rarity(ment_abund, 0)

## Hill-Shannon diversity: -ure
ure_abund <- as.numeric(as.vector(hill_abund[3,]))
rarity(ure_abund, 0)
```

## Competition

```{r Bray-Curtis}
## Abundance table
bray_abundance <- splist2presabs(keep,
                  sites.col = "SUFFIX", 
                  sp.col = "COMBINED_TYPE", 
                  keep.n = TRUE)
bray_abundance <- bray_abundance %>%
                  remove_rownames %>% 
                  column_to_rownames(var = "SUFFIX")
bray_abundance

## Dissimilarity matrix
bray_dist <- vegdist(bray_abundance, method="bray")
as.data.frame(as.matrix(bray_dist))

## Similarity matrix
bray_sim <- 1-as.matrix(bray_dist)
as.data.frame(as.matrix(bray_sim))
```


